EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-24 03_edipack2.tex
Updates from Overleaf

From a computational perspective, constructing a symmetry
sector amounts to defining a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ which relates the states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
full Fock space.
This map is typically implemented as a rank-1
integer array whose size corresponds to the dimension of the sector $D_{\SS_\vec{Q}}$$.

@@

The presence of a symmetry reduces the Hamiltonian matrix to
a block diagonal form, where each block, labeled by $\vec{Q}$, has dimension
$D_{\SS_\vec{Q}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if a
additional constraints are present.
A summary of the sector properties is provided in
Table~(\ref{TabSector}). 
\begin{table}%[ht]
  \label{TabSector}
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
\end{tabularx}
\end{center}
\caption{
  {\bf Quantum Numbers Table}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem. To each value
(first column) correspond different quantum numbers (second column).
The third column report the dimension of the symmetry sector for a
given realization of the quantum numbers.}
\end{table}

@@

Alternatively, setting {\tt ED\_READ\_UMATRIX=T} and {\tt ED\_USE\_KANAMORI=F}, a list of two-body operators can be supplied by the user in a properly
formatted text file, whose name is specified by the input variable {\tt UMATRIX\_FILE}. The implementation of the list of interaction terms is contained in
the module {\tt ED\_PARSE\_UMATRIX}.  
The file must have the suffix {\tt <UMATRIX\_FILE>.re\-star\-t} and should be formatted
as follows: 
\begin{lstlisting}[style=mybash,numbers=none]
NORB BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the numer of orbitals $N_\a$ and each line
defines a single two-body term of the form 
$U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. Note that the
coefficient indices are swapped compared to the standard two-body
operator formulation \cite{??}.  
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

\begin{lstlisting}[style=fstyle,numbers=none]    
!Single element of the list: contains a single eigen-solution.
!Twin states are just pointers, no memory.
!double or cmplx vector are automatically MPI distributed.
type sparse_estate
   integer                             :: sector        !Symmetry sector index
   real(8)                             :: e             !eigenvalue: ordering key
   real(8),dimension(:),allocatable    :: dvec          !double  eigen-vector
   complex(8),dimension(:),allocatable :: cvec          !complex eigen-vector
   logical                             :: itwin=.false. !twin sector logical label
   type(sparse_estate),pointer         :: twin=>null()  !link to twin state in the list
   type(sparse_estate),pointer         :: next=>null()  !List next pointer
end type sparse_estate
!Ordered  list storing the lower part of the Hamiltonian spectrum: 
type sparse_espace
   integer                     :: size               !Current list
   real(8)                     :: emax               !Max energy  fixed by ${\color{}e^{-\beta {\tt emax}}<{\tt cutoff}}$
   real(8)                     :: emin               !Min energy:the groundstate energy
   logical                     :: status=.false.     !Allocation status
   type(sparse_estate),pointer :: root=>null()       !Root of the linked list
end type sparse_espace

@@

The diagonal spin, charge and pair susceptibility terms
$\chi^{S^z}_{\a\a}$, $\chi^N_{\a\a}$ and $\chi^\Delta_{\a\a}$
are constructed using the operators $S^z_\a = \sum_{\s\s'} c^+_{\a\s}
\tau^z_{\s\s'} c_{\a\s'}$, 
$N_\a = \sum_{\s\s'} c^+_{\a\s} \tau^0_{\s\s'} c_{\a\s'}$
and $\Delta_\a = c_{\a\dw}c_{\a\up}$,
where $\tau^{a=0,x,y,z}$ are the Pauli matrices.
Off-diagonal terms are  evaluated using the operators
$(S^z_\a + S^z_\b)$, $(N_\a + N_\b)$ or $(\Delta_\a + \Delta_\b)$. 
The Excitonic susceptibilities, $\chi^\vec{T}_{\a\b}$, are defined with 
respect to the vector operator: $T^i_{\a\b} = \sum_{\s\s'}
c^+_{\a\s}\tau^i_{\s\s'}c_{\b\s'}$, where $i=0$ represents the
spin-singlet exciton, and $i=x,y,z$  correspond to the spin-triplet exciton.

>>>>

From a computational perspective, constructing a symmetry
sector amounts to defining a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ which relates the states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
full Fock space.
This map is typically implemented as a rank-1
integer array whose size corresponds to the dimension of the sector $D_{SS_{\vec{Q}}}$.

@@

The presence of a symmetry reduces the Hamiltonian matrix to
a block diagonal form, where each block, labeled by $\vec{Q}$, has dimension
$D_{\SS_{\vec{Q}}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if a
additional constraints are present.
A summary of the sector properties is provided in
Table~(\ref{TabSector}). 
\begin{table}%[ht]
  \label{TabSector}
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
\end{tabularx}
\end{center}
\caption{
  {\bf Quantum Numbers Table}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem. To each value
(first column) correspond different quantum numbers (second column).
The third column report the dimension of the symmetry sector for a
given realization of the quantum numbers.}
\end{table}

@@

Alternatively, setting {\tt ED\_READ\_UMATRIX=T} and {\tt ED\_USE\_KANAMORI=F}, a list of two-body operators can be supplied by the user in a properly
formatted text file, whose name is specified by the input variable {\tt UMATRIX\_FILE}. The implementation of the list of interaction terms is contained in
the module {\tt ED\_PARSE\_UMATRIX}.  
The file must have the suffix {\tt <UMATRIX\_FILE>.re\-star\-t} and should be formatted
as follows: 
\begin{lstlisting}[style=mybash,numbers=none]
NORB BANDS
i1 j1 k1 l1 U_i1j1k1l1
i2 j2 k2 l2 U_i2j2k2l2
...
\end{lstlisting}
where {\tt NORB} is the numer of orbitals $N_\a$ and each line
defines a single two-body term of the form 
$U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}$. Note that the
coefficient indices are swapped compared to the standard two-body
operator formulation $$
\frac{1}{2}\sum_{i,j,k,l} c^{\dagger}_i c^{\dagger}_j U_{ijkl} c_l c_k
$$ 
Empty lines and lines starting with {\tt \#, \%, !} are ignored.

@@

\begin{lstlisting}[style=fstyle,numbers=none]    
!Single element of the list: contains a single eigen-solution.
!Twin states are just pointers, no memory.
!double or cmplx vector are automatically MPI distributed.
type sparse_estate
   integer                             :: sector        !Symmetry sector index
   real(8)                             :: e             !eigenvalue: ordering key
   real(8),dimension(:),allocatable    :: dvec          !double  eigen-vector
   complex(8),dimension(:),allocatable :: cvec          !complex eigen-vector
   logical                             :: itwin=.false. !twin sector logical label
   type(sparse_estate),pointer         :: twin=>null()  !link to twin state in the list
   type(sparse_estate),pointer         :: next=>null()  !List next pointer
end type sparse_estate
!Ordered  list storing the lower part of the Hamiltonian spectrum: 
type sparse_espace
   integer                     :: size               !Current list
   real(8)                     :: emax               !Max energy  fixed by ${e^{-\beta {\tt emax}}<{\tt cutoff}}$
   real(8)                     :: emin               !Min energy:the groundstate energy
   logical                     :: status=.false.     !Allocation status
   type(sparse_estate),pointer :: root=>null()       !Root of the linked list
end type sparse_espace

@@

The diagonal spin, charge and pair susceptibility terms
$\chi^{S^z}_{\a\a}$, $\chi^N_{\a\a}$ and $\chi^\Delta_{\a\a}$
are constructed using the operators $S^z_\a = \sum_{\s\s'} c^+_{\a\s}
\tau^z_{\s\s'} c_{\a\s'}$, 
$N_\a = \sum_{\s\s'} c^+_{\a\s} \tau^0_{\s\s'} c_{\a\s'}$
and $\Delta_\a = c_{\a\dw}c_{\a\up}$,
where $\tau^{a=0,x,y,z}$ are the Pauli matrices.
Off-diagonal terms are  evaluated using the operators
$(S^z_\a + S^z_\b)$, $(N_\a + N_\b)$ or $(\Delta_\a + \Delta_\b)$. 
The Excitonic susceptibilities, $\chi^{\vec{T}}_{\a\b}$, are defined with 
respect to the vector operator: $T^i_{\a\b} = \sum_{\s\s'}
c^+_{\a\s}\tau^i_{\s\s'}c_{\b\s'}$, where $i=0$ represents the
spin-singlet exciton, and $i=x,y,z$  correspond to the spin-triplet exciton.
